---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# shoelati

#insert my hex sticker

<!-- badges: start -->
<!-- badges: end -->

shoelati uses data on shoes collected from a particular region to understand the relationships between the variables in the data and how they influence one another. This package makes visualizations which are more accessible, conclusive, and fun. Specifically, it uses package uses packages such as viridis to make the visualizations more accessible for all individuals, including those who suffer from colorblindness. It also performs commmon summary functions to find trends in numeric variables after grouping or filtering by numeric or categorical variables. Finally, it demonstrates how we can use geoms such as emojis to make more fun graphs, while also uncovering how general shoe types can cluster in particular variables.  

## Installation

You can install the development version of shoelati from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("harvard-stat108s23/shoelati")
```

##Exploring the data
Let's first explore the different variables in the data and their diverse range of inputs.

```{r}
library(shoelati)
library(tidyverse)

```

```{r sum-shoelati}
summary(shoelati)
```

## Examples

We can wrangle the data using unique variables to perform summary functions: 

```{r sum-table}
shoelati |>
  group_by(brand) |>
  summarize(mean = mean(price_usd),
            sd = sd(price_usd),
            median = median(price_usd),
            max = max(price_usd),
            min = min(price_usd),
            count = n())

```


We can also create meaningful data vidualizations:

```{r price-brand}
library(viridis)

ggplot(data = shoelati, mapping = aes(x = price_usd, y = brand, color = brand)) +
  geom_violin() +
  guides(color = "none") +
  scale_color_viridis(discrete = TRUE) +
  labs(title = "How does Price Distribution vary according to Brand",
       x = "Price (USD)",
       y = "Brand",
       caption = "Each color represents a different brand.") +
  theme(text = element_text(size = 15))

```



You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/v1/examples>.


In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.
